Evolution of pre-copulatory and post-copulatory strategies of inbreeding avoidance

### 2017-05-02 ms_pre_post.Rmd
Edit through Introduction

<!---

I'm confused by Ryan's point:

''It is 'required' in this sense but could a reviewer flag up that other forms of post-cop inbreeding avoidance could evolve without the need of polyandry?

I'm thinking in terms of sperm rejection/embryo inviability or something along those lines (such like you mention in the discussion), since females could then mate with a different male for subsequent events.  Very pedantic I know, but maybe a word or two of clarity re: polyandry is required for this specific type of post-cop inbreeding avoidance could cut this potential criticism off right away.''

If females mate with a different male, then it must be during a different bout of reproduction (else it's not polyandry as we define it). Perhaps what is meant is that inviability in one event effectively kills any parental investment that could affect future reproductive bouts? To me, it isn't really a different 'bout' if there is lingering parental investment that could affect current reproduction, but I guess I can see how a very pedantic reviewer could get confused. To make it completely air-tight (though a bit dense), I've added ''through fertilisation biased toward non-relatives''.

From the perspective of a breeding female, inbreeding avoidance might be achieved by avoiding mating with related males (i.e., pre-copulatory inbreeding avoidance), or by biasing fertilisation towards unrelated males after mating has occurred (i.e., post-copulatory inbreeding avoidance). Evolution of post-copulatory inbreeding avoidance through fertilisation biased toward non-relatives requires that females express some degree of polyandry, defined as mating with multiple males during a single reproductive bout. Such polyandry might initially be under selection and evolve if it allows for pre-copulatory inbreeding avoidance when females have the opportunity to mate with additional unrelated males following initial mating with a relative, assuming sufficiently strong inbreeding depression [@Duthie]. Polyandry can thereby facilitate pre-copulatory inbreeding avoidance even without any post-copulatory female choice or otherwise biased fertilisation among sperm provided by females' mates (i.e., under conditions of a 'fair raffle'). Overall, polyandry can therefore simultaneously allow females to mate with less closely related males and create the opportunity for further inbreeding avoidance enacted through active post-copulatory choice. Polyandry has consequently been widely hypothesised to evolve as an adaptation to allow females to avoid inbreeding and thereby increase offspring fitness [@Zeh1997; @Jennions2000; @Tregenza2002]. Selection on polyandry to increase offspring fitness could explain its evolution in cases where multiple mating does not increase female reproductive success, or even reduces reproductive success and therefore imposes direct costs on polyandrous females. However, despite such widely-invoked hypotheses regarding the role of inbreeding depression on evolving strategies for pre-copulatory and post-copulatory inbreeding avoidance and hence polyandry, there is surprisingly little theory or modelling that critically examines the conditions under which pre-copulatory or post-copulatory inbreeding avoidance, or both strategies, are predicted to evolve, or that examines how the evolution of such strategies of inbreeding avoidance might feed back to affect underlying evolution of polyandry.

Despite the paucity of theory, numerous empirical studies on diverse species have tested for, and in some cases found evidence of, female inbreeding avoidance in systems where polyandry also occurs [@Tregenza2002; @Varian-Ramos2012; @Kingma2013; @Arct2015; but see @Reid2015a]. However, relatively few studies have determined whether inbreeding avoidance is enacted through pre-copulatory or post-copulatory mechanisms. Pre-copulatory inbreeding avoidance has been reported in diverse species, spanning sweet potato weevils [*Cylas formicarius*, @Kuriwada2011], purple-crowned fairy-wrens [*Malurus coronatus*, @Kuriwada2011], and squinting bush brown butterflies [*Bicyclus anynana*, @Fischer2015], while post-copulatory inbreeding avoidance has been reported in red junglefowl [*Gallus gallus*, @Pizzari2004] and crickets [*Teleogryllus oceanicus*, @Simmons2006; *Gryllus bimaculatus*, @Bretman2009]. Evidence for both pre-copulatory and post-copulatory inbreeding avoidance is available across different studies of Trinidadian guppies [*Poecilia reticulata*; @Gasparini2011; @Daniel2015] and house mice [*Mus domesticus*; @Potts1991; @Firman2015]. Meanwhile, @Liu2014 found evidence of pre-copulatory but not post-copulatory inbreeding avoidance within a single study of cabbage beetles (*Colaphellus bowringi*). However, @Ala-Honkola2011 and @Tan2012 found no evidence for pre-copulatory or post-copulatory inbreeding avoidance in fruit flies (*Drosophila melanogaster*), respectively, and Reid et al. [-@Reid2014; -@Reid2015] found no net inbreeding avoidance in song sparrows (*Melospiza melodia*) despite strong inbreeding depression, suggesting neither pre-copulatory or post-copulatory inbreeding avoidance. Taken together, these studies demonstrate that diverse combinations of pre-copulatory and post-copulatory inbreeding avoidance occur in nature. However, there is as yet no theory that predicts what combinations of pre-copulatory and post-copulatory inbreeding avoidance and and associated polyandry should be favoured by selection when all combinations of phenotypes can evolve. Consequently, there is no theory that allows the diversity of observed pre-copulatory and post-copulatory inbreeding avoidance to be interpreted, and no clear hypotheses that could be tested through future empirical studies of individual systems or subsequent comparative analyses.

@Duthie used a genetically-explicit individual-based model to examine the conditions under which polyandry is predicted to evolve solely due to selection stemming from pre-copulatory inbreeding avoidance. Net selection for polyandry occurred only under highly restricted conditions, requiring that direct negative selection (i.e., 'costs') on polyandry was weak, that very few males were available for a female's initial mate choice but many were available for additional mate choice(s), or that polyandry was conditionally expressed when a focal female was related to her initial mate [@Duthie]. Without these conditions, and without the possibility for post-copulatory female choice, increasing polyandry tended to increase rather than decrease the overall degree of inbreeding that females experienced, and hence reduce offspring fitness. This occurred because, once pre-copulatory inbreeding avoidance evolved, polyandrous females had already chosen available unrelated males as their initial mates. Their additional mates, chosen from the remaining male population, were therefore increasingly likely to include relatives. Evolution of polyandry purely to facilitate pre-copulatory inbreeding avoidance, and hence the resulting degree of inbreeding avoidance, was consequently restricted [@Duthie]. If post-copulatory inbreeding avoidance could also evolve, polyandrous females would have the opportunity to bias paternity towards unrelated males within their set of chosen mates. Evolution of post-copulatory inbreeding avoidance might consequently reduce the cost of polyandry stemming from the accumulation of related mates across multiple matings, potentially facilitating evolution of polyandry to avoid inbreeding under broader conditions, and driving further evolution of pre-copulatory or post-copulatory mate choice strategies.

However, if polyandry and pre-copulatory and post-copulatory inbreeding avoidance can all evolve, the long-term dynamics of these three reproductive strategies become difficult to predict. Strong inbreeding depression might drive initial evolution of both pre-copulatory and post-copulatory inbreeding avoidance and associated polyandry. However, the co-occurence of pre-copulatory and post-copulatory inbreeding avoidance might engender some degree of 'degeneracy', defined as a phenomenon by which different elements of a system result in identical outputs [i.e., @Edelman2001]. Consequently, if evolution of polyandry and post-copulatory inbreeding avoidance renders pre-copulatory inbreeding avoidance functionally redundant, or vice versa, then only one inbreeding avoidance strategy might be maintained in the long-term.

@@

Ultimately, the absolute and relative frequencies of alleles underlying pre-copulatory and post-copulatory inbreeding avoidance and polyandry will depend not only on the magnitudes of positive indirect selection stemming from reduced inbreeding depression in females' offspring, but also on the magnitudes of direct negative selection on resulting phenotypes (i.e., the direct fitness costs of expressing each reproductive strategy). Polyandrous females can pay costs of mate searching or mating, for example, because these activities increase predation risk [e.g., @Rowe1988; @Ronkainen1994; @Koga1998]. Females that express pre-copulatory choice and consequent reluctance to mate with relatives might pay costs of increased risk of complete mating failure or harm caused by sexual conflict [@Rowe1994; @Kokko2013]. Females that express post-copulatory choice might pay initial energetic costs associated with developing the physiological or biochemical mechanisms needed to store sperm and successfully bias fertilisation [e.g., @Gasparini2011; @Tuni2013; @Fitzpatrick2014b]. If the relative costs of pre-copulatory and post-copulatory inbreeding avoidance differ, then alleles underlying the less costly strategy might become fixed over generations, while alleles underlying the more costly inbreeding avoidance strategy might become extinct if their effects are both costly and made redundant by the less costly strategy. However, such trajectories will also depend on the indirect benefits of each strategy, which might also differ and vary in relation to the evolution of the other strategy. Dynamic models that track the frequencies of alleles underlying multiple reproductive strategies that might influence each others evolution across generations, and that are enacted among relatives that themselves result from reproductive strategies and inbreeding depression expressed in previous generations, are consequently useful to understand and predict evolutionary outcomes.

>>>>

From the perspective of a breeding female, inbreeding avoidance might be achieved by avoiding mating with related males (i.e., pre-copulatory inbreeding avoidance), or by biasing fertilisation towards unrelated males after mating has occurred (i.e., post-copulatory inbreeding avoidance). Evolution of such post-copulatory inbreeding avoidance is likely to be facilitated females express some degree of polyandry, defined as mating with multiple males during a single reproductive bout [but see @Dougherty2016]. Such polyandry might itself be under selection and evolve if it allows females to mate with additional unrelated males following an initial mating with a relative. Polyandry can thereby facilitate pre-copulatory inbreeding avoidance even without any post-copulatory female choice or otherwise biased fertilisation among sperm provided by females' mates (i.e., under conditions of a 'fair raffle'). Overall, polyandry can therefore simultaneously allow females to mate with less closely related males and create the opportunity for further inbreeding avoidance enacted through active post-copulatory choice. Polyandry has consequently been widely hypothesised to evolve as an adaptation to allow females to avoid inbreeding and thereby increase offspring fitness [@Zeh1997; @Jennions2000; @Tregenza2002]. Resulting indirect selection on polyandry to increase offspring fitness could explain evolution of polyandry in cases where multiple mating does not increase female reproductive success, or even reduces reproductive success and therefore imposes direct costs on polyandrous females. However, despite such widely-invoked hypotheses regarding the role of inbreeding depression in driving pre-copulatory and post-copulatory strategies and polyandry, there is surprisingly little theory or modelling that critically examines the conditions under which pre-copulatory or post-copulatory inbreeding avoidance, or both strategies, are predicted to evolve, or that examines how the evolution of such strategies of inbreeding avoidance might feed back to affect underlying evolution of polyandry.

<!--- I really don't think the connection between 'resulting indirect selection' and *offspring* rather than *parent* fitness is going to be made easily by a lot of readers. You've thought a lot about this, so I think we need to be careful for readers who are still not up to date on the selection-on-inbreeding versus selection-on-being-inbred distinction. --->

Despite the paucity of theory, numerous empirical studies on diverse species have tested for, and in some cases found evidence of, female inbreeding avoidance in systems where polyandry also occurs [@Tregenza2002; @Varian-Ramos2012; @Kingma2013; @Arct2015; but see @Reid2015a]. However, relatively few studies have determined whether inbreeding avoidance is enacted through pre-copulatory or post-copulatory mechanisms. Pre-copulatory inbreeding avoidance has been reported in diverse species, spanning sweet potato weevils [*Cylas formicarius*, @Kuriwada2011], purple-crowned fairy-wrens [*Malurus coronatus*, @Kuriwada2011], and squinting bush brown butterflies [*Bicyclus anynana*, @Fischer2015], while post-copulatory inbreeding avoidance has been reported in red junglefowl [*Gallus gallus*, @Pizzari2004] and crickets [*Teleogryllus oceanicus*, @Simmons2006; *Gryllus bimaculatus*, @Bretman2009]. Evidence for both pre-copulatory and post-copulatory inbreeding avoidance is available across different studies of Trinidadian guppies [*Poecilia reticulata*; @Gasparini2011; @Daniel2015] and house mice [*Mus domesticus*; @Potts1991; @Firman2015]. Meanwhile, @Liu2014 found evidence of pre-copulatory but not post-copulatory inbreeding avoidance within a single study of cabbage beetles (*Colaphellus bowringi*). However, @Ala-Honkola2011 and @Tan2012 found no evidence for pre-copulatory or post-copulatory inbreeding avoidance in fruit flies (*Drosophila melanogaster*), respectively, and Reid et al. [-@Reid2014; -@Reid2015] found no net inbreeding avoidance in song sparrows (*Melospiza melodia*) despite strong inbreeding depression and opportunities for both pre-copulatory and post-copulatory inbreeding avoidance. Taken together, these studies demonstrate that diverse combinations of pre-copulatory and post-copulatory inbreeding avoidance, or lack thereof, occur in nature. However, there is as yet no theory that predicts what combinations of pre-copulatory and post-copulatory inbreeding avoidance and associated polyandry should be favoured by selection when all can evolve. Consequently, there is no theory that allows the diversity of observed pre-copulatory and post-copulatory inbreeding avoidance strategies to be interpreted, and no clear hypotheses that could be tested through future empirical studies of individual systems or subsequent comparative analyses.

@Duthie used a genetically-explicit individual-based model to examine the conditions under which polyandry is predicted to evolve due to selection stemming from pre-copulatory inbreeding avoidance in the absence of any post-copulatory inbreeding strategy. Net selection for polyandry occurred only under highly restricted conditions, requiring that direct negative selection (i.e., 'costs') on polyandry was weak, that very few males were available for a female's initial mate choice but many were available for additional mate choice(s), or that polyandry was conditionally expressed when a focal female was related to her initial mate [@Duthie]. Without these conditions, increasing polyandry tended to increase rather than decrease the overall degree of inbreeding that females experienced, and hence reduce offspring fitness. This occurred because, once pre-copulatory inbreeding avoidance evolved, polyandrous females had already chosen available unrelated males as their initial mates. Their additional mates, chosen from the remaining male population, were therefore increasingly likely to include relatives. Evolution of polyandry purely to facilitate pre-copulatory inbreeding avoidance, and hence the resulting degree of inbreeding avoidance, was consequently restricted [@Duthie].

However, if post-copulatory inbreeding avoidance could also evolve, polyandrous females would have the opportunity to bias paternity towards unrelated males within their set of chosen mates. Evolution of post-copulatory inbreeding avoidance might consequently reduce the cost of polyandry stemming from the accumulation of related mates across multiple matings, potentially facilitating evolution of polyandry to avoid inbreeding under broader conditions, and driving further evolution of pre-copulatory or post-copulatory mate choice strategies. Subsequently, if polyandry and pre-copulatory and post-copulatory inbreeding avoidance can all evolve, the long-term dynamics of these three reproductive strategies become difficult to predict. Strong inbreeding depression might drive initial evolution of both pre-copulatory and post-copulatory inbreeding avoidance and associated polyandry. However, the co-occurrence of pre-copulatory and post-copulatory inbreeding avoidance might engender some degree of 'degeneracy', defined as a phenomenon by which different elements of a system result in identical outputs [i.e., @Edelman2001]. Consequently, if evolution of polyandry and post-copulatory inbreeding avoidance renders pre-copulatory inbreeding avoidance functionally redundant, or vice versa, then only one inbreeding avoidance strategy might be maintained in the long-term.

@@

Ultimately, the absolute and relative frequencies of alleles underlying pre-copulatory and post-copulatory inbreeding avoidance and polyandry will depend not only on the magnitudes of positive indirect selection stemming from reduced inbreeding depression in females' offspring, but also on the magnitudes of direct negative selection on resulting phenotypes (i.e., the direct fitness costs of expressing each reproductive strategy). Polyandrous females can pay costs of mate searching or mating, for example, because these activities increase predation risk [e.g., @Rowe1988; @Ronkainen1994; @Koga1998]. Females that express pre-copulatory choice and consequent reluctance to mate with relatives might pay costs of increased risk of complete mating failure or harm caused by sexual conflict [@Rowe1994; @Kokko2013]. Females that express post-copulatory choice might pay initial energetic costs associated with developing the physiological or biochemical mechanisms needed to store sperm and successfully bias fertilisation [e.g., @Gasparini2011; @Tuni2013; @Fitzpatrick2014b]. If the relative costs of pre-copulatory and post-copulatory inbreeding avoidance differ, then alleles underlying the less costly strategy might become fixed over generations, while alleles underlying the more costly strategy might become extinct, especially if their effects are made redundant by evolution of the less costly strategy. Evolutionary dynamics will therefore depend on the benefits of each strategy, which might also differ and vary in relation to the evolution of the other strategy. Dynamic models that track the frequencies of alleles underlying multiple reproductive strategies that might influence each others evolution across generations, and that are enacted among relatives that themselves result from reproductive strategies and inbreeding depression expressed in previous generations, are consequently useful to understand and predict evolutionary outcomes.
